Recruiting Chapter

### 2021-06-02 text/front_end_june_2021.Rmd
draft changes front end text

---
title: "Recruiting Chapter"
author: 
  - Ozan Jaquette
  - Crystal Han
  - Irma Castaneda
bibliography: ./other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \floatsetup{capposition=top}
---

@@

<!-- 
A pair of papers by Stevens provide interesting statements about the functions of higher education and the behavior of colleges and universities
Colleges and universities visit high schools that enroll desirable prospects, while high schools accept visits from colleges and universities that offer enrollment opportunities valued by some segment of the class. Drawing from a pair of papers by Mitchell Stevens [@RN2377; @RN4517], we argue that analysis of which high schools receive visits from which colleges and univeristies provides an opportunity how K-12 and postsecondary organizations form relationships that serve social reproduction.

This chapter presents a social network analysis of off-campus recruiting visits made by US colleges and universities to US private high schools. Colleges and universities visit high schools that enroll desirable prospects, while high schools accept visits from colleges and universities that offer enrollment opportunities valued by some segment of the class. An off-campus recruiting visit is an indicator of a relationship between a college and a school that can be analyzed using standard tools of social network analysis. We analyze visits made in 2017 from a sample of 25 selective private colleges and universities and a sample of 17 public research universities.  Analyses investigate which types of private high schools receive visits from which types of colleges and universities, emphasizing comparisons within and between public and private postsecondary institutions.

The motivation for these analyses is grounded within the sociology of higher education, particularly two theoretical and empirical syntheses developed by Mitchell Stevens and colleagues [@RN2377; @RN4517]. @RN2377 introduces metaphors that describe how sociology has conceived the functions of higher education. The stratification tradition conceives of higher education as a "sieve" -- an instrument that strains and separates -- that allocates people to occupational and social positions. This metaphor builds on "Max Weber's core insight that education has a dual character" [@RN2377, p. 129], on one hand, serving social mobility by redistributing access to privileged positions and, on the other, serving social reproduction by granting privileged families disproportionate access to universities that confer privileged positions. The institutional theory tradition in sociology conceives of higher education as a "hub" that connects, pollinates, and legitimizes "prominent institutional sectors of modern societies" (e.g., professions, the philanthropic sector, corporate sector).  As a hub, higher education is "simultaneously a powerful and fragile social institution. On the one hand, higher education [is powerful because it] connects and reciprocally blesses various forms of privilege. Elite groups (e.g., high-status families, professions, politicians, scientists, intellectuals) and those aspiring to be a part of them use higher education to certify their legitimacy" [@RN2377, p. 137]. On the other hand, higher education is fragile because it depends on "patronage as a primary source of revenue."

@RN4517 introduces three alternative metaphors to describe the relationship between higher education and the American political economy. The "association" conceptualization, observed by @RN485, views colleges as voluntary (i.e., non-governmental) organizations founded by voluntary associations -- principally religious congregations and local business interests -- that rely "on the collective contributions and goodwill of their memberships" [@RN2377, p. 126]. Second, the "national service" configuration views higher education as a public-good engine of economic development and nation building and thus worthy of investment from federal and state governments. Third, "market" configuration conceives of higher education as a private good, suggesting that colleges should survive by competing for customers rather than rely on public funding. Each of these configurations is more or less associated with particular historical epochs and particular types of colleges and universities. For example, the association configuration defined US higher education through the 19th century, but remains salient for private colleges and universities.

<!-- 
The national service configuration is associated associated with massive public investment in higher education in the decades following World War II, while the market configuration is associated with declining state support for public universities since the 1980s.

These metaphors -- which describe higher education simultaneously serving social mobility and social reproduction and depending on patronage from the government, from voluntary associations, and private households -- are useful for comparing public flagship universities to selective private colleges and universities
-->

These metaphors are useful for comparing public flagship universities to selective private colleges and universities. Selective private colleges and universities depend on patronage from affluent households (market configuration) and philanthropic interests (association configuration) and primarily serve social reproduction rather than social mobility. Historically, public flagship universities depended on government funding associated with the national service configuration and served social mobility rather than reproduction. Since the 1980s, reductions in public funding associated with the emergence of the market configuration encourage public flagship universities to prioritize patronage from affluent households and philanthropic interests. As suggested by the adage, "he who pays the piper calls the tune," reliance on affluent private-sector interests increases the likelihood that public flagship universities -- like selective private colleges and universities -- will primarily serve social reproduction rather than social mobility.
<!-- 
The extent to which higher education serves reproduction versus redistribution is sensitive to the external funding environment because universities depend on ``patronage as a primary revenue source'' [@RN2377, p. 138]. 
-->

We argue that the recruiting behavior of colleges and universities is an indicator of organizational enrollment priorities and, in turn, the extent to which enrollment priorities serve social mobility or social reproduction. However, access to concrete data on university recruiting behaviors has been a persistent barrier to empirical research. We collected data on off-campus recruiting visits made in 2017 by "scraping" URLs (e.g., "Coming to a neighborhood near you") once per week from college/university admissions websites and -- for a sample of public research universities -- by issuing public records requests. The first study from this data collection analyzed visits made by 15 public research universities [@RN4450]. The overwhelming majority of public research universities in our sample made more out-of-state visits than in-state visits, a finding antithetical to historic mission of public universities. These out-of-state visits focused on wealthy, predominantly white schools and communities, with a disproportionate number of visits to out-of-state private high schools but relatively few visits to in-state private high schools.

Building on @RN4450, this chapter analyzes off-campus recruiting visits to private high schools made in 2017 by a sample of 25 selective private colleges and universities and 17 public research universities. We utilize social network methods and emphasize comparisons within and between public and private postsecondary institutions. Analyses are informed by two broad, related questions. First, which types of private high schools receive visits from which types of colleges and universities? Second, to what extent do colleges and universities visit private high schools that share similar characteristics with the college (e.g., geographic region, religious affiliation, academic reputation, and racial composition)?

<!-- 
Our decision to compare public and private colleges/universities on the dimension of visits to private schools is motivated by the idea that recruiting from private high schools serves social reproduction rather than social mobility.

Our decision to compare public and private colleges/universities on the dimension of visits to private schools is motivated by the idea that recruiting from private high schools serves social reproduction rather than social mobility. Thus, visits to private schools are antithetical to the historic mission of public universities, especially given that the vast majority of visits are to out-of-state private schools. By contrast, recruiting from private high schools is consistent with the mission of selective private colleges and universities. From this perspective, visits to private schools by public research universities is one example of "privatization." A weakness of the empirical literature on privatization in higher education is that few studies public and private colleges/universities with respect to a behavior associated with privatization [@RN4543]. By comparing public and private colleges/universities on the behavioral dimension of recruiting visits to private schools, we hope to develop new insights about the privatization of public higher education.

Our decision to analyze visits to private schools is motivated by the idea that recruiting from private schools largely serves social reproduction<!-- [CITE A SOCIOLOGIST]-->. Thus, visits to private schools are antithetical to the historic "social mobility mission" of public universities [@RN2269], particularly since most private school visits by public research universities are to out-of-state schools [@RN4450]. By contrast, recruiting from private high schools is consistent with the missions of selective private colleges and universities. From this perspective, private school visits by public research universities is an example of "privatization" that serves social reproduction.  A weakness of the empirical literature on privatization in higher education is that few studies public and private colleges/universities with respect to a behavior associated with privatization [@RN4543]. By comparing public and private colleges/universities on the behavioral dimension of recruiting visits to private schools, we hope to develop new insights about the privatization of public higher education.

Finally, although private school students are more likely than public school students to come from affluent, college-educated households, the population of US private schools contains tremendous variation in religious affiliation, academic reputation, racial and socioeconomic composition, amongst other dimensions [@RN4504]. Therefore, focusing solely on visits to private schools enables us to develop a nuanced analysis of which types of private high schools receive visits from which types of colleges and universities. These nuances would be difficult to observe in an analysis that also included visits to public high schools.

<!-- 
We use social network analysis (SNA) to analyze off-campus recruiting visits. The motivation for using SNA is that SNA privileges relationships between actorsQualitative scholarship on private high schools finds that college access is substantially determined by the quality of the relationship between a particular college and a particular private high school [@RN4321; @RN4407]. An off-campus recruiting visit is an indicator of a relationship between a college and a school that can be analyzed using standard tools of SNA.
-->

# What Do We Know About Off-Campus Recruiting?

<!-- 
This section provides context for our analyses by, first, situating off-campus visits within the broader set of marketing and recruiting interventions. Second, we review what is known about off-campus recruiting from market research and from academic scholarship, emphasizing scholarship that examines the relationship between private high schools and colleges/universities.
-->

@@

<!--
TEXT BRIEFLY DESCRIBING FINDINGS FROM DESCRIPTIVE TABLES.... Table/Figure X demonstrates the disproportionate number of visits to private high schools.

WHAT TO SHOW IN TABLES/FIGURES:
- Table/figure of number of visits by visit type and in/out-of-state
- Table/figure showing disproportionate number of visits to private high schools.
-->

# Network Methods and Data

We utilize network methods rather than conventional descriptive statistics for two, related reasons. First, social network analysis privileges the relationship between actors. The discussion of Stevens and Khan suggests that for private school students, "it is not just the quality of the students that gets them into college but the quality of the relationship between elite high schools and colleges" [@RN4407, p. 175]. Second, a visit from a college/university admissions counselor to a high school can be conceived as a "network tie" connecting two actors in a social network.

We briefly define essential social network analysis concepts and their operationalization in our analyses.  A social network consists of a set of actors -- referred to as "vertices" -- and the connections -- referred to as "edges" between these actors.  Whereas "one-mode" networks consist of vertices of the same "type" (e.g., in a publication network each vertex is an author), "two-mode" networks consist of vertices associated with one type of actor/entity having connections to vertices of another type. For example, an actor-movie network consists of actors (mode 1) who appear in movies (mode 2), and an actor shares an edge with a movie if the actor appears in the movie. <!--The social network analyzed in this chapter is a two-mode network, where vertices consists of colleges/universities (mode 1) and high schools (mode 2), and an edge is defined as a visit from a college/university to a private high school [DELETE THIS SENTENCE OR THE REDUNDANT ONE IN BELOW PARAGRAPH].-->

Our analyses draw from analyses of "corporate board-director" networks because these are the most commonly analyzed two-mode networks in the social networks literature [e.g., @RN4663]. Board-director networks consist of directors (mode 1) who serve on organizational boards (mode 2). In our school-college two-mode network, the vertices are private high schools (mode 1) and colleges/universities (mode 2), and an edge is defined as an off-campus recruiting visits made between the two (e.g., high school $i$ shares an edge with college $j$ if high school $i$ receives at least one visit from college $j$). Visits can only occur between a school-college pair (not a school-school pair or a college-college pair). This (weighted) network can be represented as a school-by-college matrix (e.g., a $500 \times 40$ matrix if our network contains 500 high schools and 40 colleges/universities) in which matrix cell $a_{i,j}$ identifies the number of visits that high school $i$ received from college/university $j$. <!--[SHOW AN EXCERPT OF THE TWO MODE ADJACENY MATRIX, LIKE 4 BY 3 OR SOMETHING] [DEFINE WEIGHTED VS. UNWEIGHTED SOMEWHERE]-->

## Two-mode vs. One-mode networks

Two-mode networks are often analyzed as one-mode networks [@RN2457]. @RN4663 states that "overlapping groups such as boards of directors form a two-mode membership network in which one can conceive of directors as nodes [vertices] connected by a tie of common board membership, or boards as nodes connected by a tie of one or more shared directors" (p. XXX). Thus, a two-mode board-director network can be transformed into two one-mode networks. A weighted one-mode director network can be created as follows: create a director-by-director matrix (e.g., a $100 \times 100$ matrix if there are 100 directors in the network); two directors share an edge if they serve together on a board; matrix cell $a_{i,j}$ identifies the number of boards that include both director $i$ and director $j$ (e.g., $a_{i,j}=0$ if directors $i$ and $j$ do not serve on any of the same boards and $a_{i,j}=3$ if directors $i$ and $j$ serve on three boards together). Similarly, a one-mode board network is created based on a board-by-board matrix in which two boards share an edge if a director sits on both boards and matrix cell $a_{i,j}$ identifies the number of directors that serve on both board $i$ and board $j$. Following this approach, we create a one-mode high school network in which two private high schools share an edge if they both receive a visit by the same college/university. Cell $s_{i,j}$ of the school-by-school matrix $S$ indicates the number of colleges/universities that visited both school $i$ and school $j$<!-- [SHOW EXCERPT OF ONE-MODE SCHOOL BY SCHOOL MATRIX, LIKE 3 BY 3 OR SOMETHING]-->. Similarly, we create a one-mode college/university network (herein the college network) in which two colleges share an edge if both colleges visited the same high school. Cell $c_{i,j}$ of the college-by-college matrix $C$ indicates the number of private high schools that received at least one visit from both college $i$ and college $j$<!-- [SHOW EXCERPT OF ONE-MODE COLLEGE BY COLLEGE MATRIX, LIKE 3 BY 3 OR SOMETHING]-->.

## Ego networks

A limitation of our data collection and analysis is that it is based on a convenience sample of 42 colleges and universities, as described above. Thus, for each college and university in our sample, we know all the private high schools they visited (assuming no measurement error). However, for each private high school in our sample, we only know visits they received from the set of 42 colleges/universities in our sample. Given this limitation, it may be useful to think of our data collection as resulting in 42 "ego networks." An "ego" is a single "focal" node (e.g., The University of Notre Dame). An ego network consists "of a focal node ('ego') and the nodes to whom ego is directly connected to (these are called 'alters') plus the ties, if any, among the alters" (p. XXX)<!-- [from http://www.analytictech.com/networks/egonet.htm#:~:text=Ego%20networks%20consist%20of%20a,form%20The%20human%20social%20network ]-->. For example, the University of Notre Dame ego network consists of all private high schools that received at least one visit from Notre Dame. Additionally, for each high school visited by Notre Dame, the ego network may include all colleges/universities in our sample that also visited the high school.

When analyzing ego networks, researchers often investigate the extent of "homophily" versus "heterophily" [@RN1722]. Homophily is the idea that two vertices are more likely to be connected if they have similar characteristics (e.g., two people are more likely to be friends on Facebook if they have similar political ideologies). Heterophily, the opposite of homophily, is when pairs of connected vertices have different characteristics from one another. In our analyses, we investigate the extent to which ego network ties are homophilous or heterophilous, focusing on the vertex characteristics of geographic region, religious affiliation, racial composition, and academic ranking.

Figure X shows the University of Notre Dame ego network, with separate panels for vertex characteristics (academic reputation, geography, religious affiliation, and racial composition) and with vertex shapes (?colors?) to identify values of vertex characteristics [INSERT FIGURE]. Visually, we can see that ....[DESCRIBE BIG TAKEAWAYS].

Figure X augments the Notre Dame ego network by including the private colleges/universities in our sample that visited private high schools visited by Notre Dame [INSERT FIGURE]. ADD BRIEF TEXT ABOUT FIGURE?

## Community detection

Many social network analyses utilize data-driven "community detection" algorithms to categorize vertices into a small number of groups (or cluster) based purely on patterns of network ties. Generally, groups are chosen to maximize the number of within-group ties while minimizing ties between members of different groups. We performed hierarchical cluster analysis to allocate colleges and universities into clusters based on their private high schools visits.<!--We used XYZ community detection approach [add footnote w/ details?] to allocate private colleges and universities [? and private high schools] to clusters.-->

# Network Analyses of Off-Campus Recruiting Visits

This section presents network analyses of visits by colleges and universities to private high schools. Analyses are motivated by two, related research questions. First, which types of private high schools receive visits from which types of colleges and universities? Second, to what extent do off-campus recruiting visits by colleges and universities to private high schools exhibit homophily versus heterophily with respect to the vertex characteristics of geographic region, religious affiliation, academic reputation, and racial composition?

[PARAGRAPH ABOUT TABLE THAT DESCRIBES CHARACTERISTICS OF VISITED VS. NON-VISITED PRIVATE SCHOOLS]
Before proceeding to network analyses, Table/Figure X compares the characteristics of private high schools that received at least one recruiting visit from a college/university in our sample to those of private high schools that did not receive a visit. WHICH CHARACTERISTICS TO SHOW: geographic region; religious affiliation; academic reputation; racial composition; enrollment size.

Figure \@ref(fig:two-mode-plot) plots the two-mode network of visits to private high schools by all colleges and universities in our sample, with large vertices indicating a college/university, small verticies indicating private high schools and with vertex color determined by geographic region (Northeast, Midwest, South, West). The graph layout places vertices closer together if they are similar to one another in terms of network structure. [SAY WHAT THIS MEANS, DRAWING FROM NOTES FROM RUSS].  Figure \@ref(fig:two-mode-plot) suggests a modest relationship between geographic region and network structure. For instance, the high school vertices closest to University of Alabama and University of Arkansas tend to be located in the south. Further, universities located in the same geographic region (e.g., SUNY Stony Brook and Rutgers) tend to be closer to one another than universities located in other regions. However, this relationship is not consistent, for example, Scripps College located in CA is closest to University of Pittsburgh (PA) and Stevens Institute of Technology (NJ). Moreover, the network depicted by \@ref(fig:two-mode-plot) is too large to detect nuanced patterns of network structure.

<!--
Figure \@ref(fig:two-mode-plot) plots the two-mode network of visits to private high schools by all colleges and universities in our sample with private high schools vertices the color salmon, public research universities shown in violet, and private colleges and universities shown in green green. The graph layout places vertices closer together if they are similar to one another in terms of network structure. [SAY WHAT THIS MEANS, DRAWING FROM NOTES FROM RUSS].  Figure \@ref(fig:two-mode-plot) suggests that the recruiting visit patterns of private colleges and universities tend to be more similar to one another than they are to those of public research universities. By contrast, many public research universities are far away from one another and a subset of public researchg universities appear to have network structure similar to many of the private colleges and universities (e.g., CU Boulder, UC Berkeley, University of South Carolina )
-->

We present results in three steps: First, visits by private colleges and universities; second, visits by public research universities; and, third, comparisons of visits by public research universities to visits by private colleges and universities.

## Visits by Private Colleges and Universities

```{r, include = FALSE, eval = FALSE}
df <- readRDS('./assets/tables/table_ego_privu.RDS')
df %>% glimpse()
df %>% select(ID,University,Cluster,total) %>% 
  arrange(desc(Cluster),University) %>% print(n=30)
```

Figure \@ref(fig:plot-1mode-privu) plots the one-mode network of visits by private colleges and universities. Here, the number of ties between a pair of universities (e.g., Sewanee and Texas Christian University) equals the number of private high schools visited by both universities. The graph layout places colleges/universities closer together if they have similar network structure [MAYBE SAY WHAT THIS MEANS IF YOU HAVEN'T YET].  Additionally, vertex color is determined by hierarchical cluster analysis specified to create $k = 4$ clusters.  Although the process of placing vertices was separate from the process of determining community cluster, vertices within the same cluster tend to be closer to one another than those of different clusters.

One cluster consists of Notre Dame, Villanova, Tulane, and Southern Methodist University. Each of these four universities made a relatively high number of visits to private high schools and are relatively centrally located in Figure \@ref(fig:plot-1mode-privu).

Addressing our two motivating research questions -- which schools receive visits from which colleges/universities and the extent to which visits are more likely to occur between organizations that share certain characteristics -- requires analyses that are more granular than Figure \@ref(fig:plot-1mode-privu). Figure \@ref(fig:ego-network-figure-privu) shows characteristics of private high schools visited our sample of private colleges and universities, specifically, geographic region, religious affiliation, academic reputation, and racial composition.

Figure \@ref(fig:ego-network-figure-privu) indicates that Notre Dame, Villanova, Tulane, and Southern Methodist University made a large share of visits in the South and the Northeast. With respect to religious affiliation, Notre Dame and Villanova -- both Catholic universities -- predictably focused visits on Catholic Schools. However, Tulane and SMU also visited a higher share of Catholic schools than most other private colleges and universities. These four universities also visited a lower share of private schools ranked in the top 200 nationally by Niche.

Figure \@ref(fig:ego-network-figure-privu) was created from the ego network of each private college and university. Ego networks consist of an ego (e.g., a single college or university), the set of high schools the ego visted (one degree of separation from the ego), and the set of colleges/universities that also visited high schools visited by the ego (two degrees of separation from the ego). Figure \@ref(fig:villanova-religion) plots the ego network of Villanova University, with vertex color determined by religious affiliation.  About 50% of Villanova's visits were to Catholic schools, a percentage [?percent?] much higher than most colleges and universities shown in Figure \@ref(fig:ego-network-figure-privu), and second only to Notre Dame University. In Figure \@ref(fig:villanova-religion), Notre Dame is closer to Villanova than any other college/university in the Villanova ego network, suggesting that Villanova and Notre Dame visit substantially overlapping sets of schools. Boston College, another Catholic university, is situated a bit further away from Villanova in the clump of colleges that includes Case Western, Tulane, Sewanee, Southern Methodist, and TCU, most of which visit relatively high share of Catholic high schools.  By contrast, Villanova is far from nonsectarian liberal arts colleges (e.g., Swarthmore, Smith, Colorado College) which visit fewer Catholic high schools.

A second cluster consits of Boston College, Colorado College, Connecticut College, Emory University, Middlebury, Northwestern Oberlin, Sewanee, Texas Christian University, Tufts, and the University of Denver.  This institutions also tend to be located relatively centrally in Figure \@ref(fig:plot-1mode-privu), indicating that many of the high schools they visit are also visited by many private colleges and universities in our sample. Compared to the first cluster, these institutions have a stronger focus on nonsectarian private high schools, with the exception of Boston College, Texas Christian, and Sewanee. The also tended to visit schools with relatively strong academic reputations.

Figure \@ref(fig:emory-region) shows the ego network of Emory University, with vertex color determined by geographic region. Located in Atlanta, about 47% of Emory's visits were to schools in the South. Figure \@ref(fig:emory-region) indicates that Emory's network structure is similar to several other southern universities -- including Tulane, TCU, SMU, and Sewanee, but is dissimilar from Baylor University which made about 56% of visits in the South, but with a focus on schools with a Christian religious affiliation. Figure \@ref(fig:emory-region) also suggests that the set of schools visited by Emory overlaps substantially with several institutions located in the Northeast (e.g., Villanova, Boston College, Middlebury) and the Midwest (e.g., Case Western Reserve, Northwestern, Notre Dame).

A third cluster consists of Case Western Reserve, MaCalester, Occidental, Smith, Swarthmore, and Williams. These institutions made relatively fewer visits than institutions in the first two clusters and are located relatively peripherally in the one-mode network, as shown in Figure \@ref(fig:plot-1mode-privu).  With the exception of Case Western Reserve, these institutions are liberal arts colleges rather than research universities.  They tended to visit a high share of nonsectarian private high schools and also a high share of schools ranked in the top 200 nationally by Niche.

The final cluster consists of Baylor, Harvey Mudd, Scripps, and Stevens Institute of Technology. Figure \@ref(fig:plot-1mode-privu) indicates that all four institutions are rather peripheral to the network. Harvey Mudd, Scripps, and Stevens made relatively few visits to private high schools. By contrast, Baylor visited a fundamentally different set of schools than other private colleges and universities in that Baylor visited a large number of Christian private schools that were not visited by other private colleges and universities in the sample. Overall, these four institutions form a group more because they are peripheral to the overall network as opposed to having similar network structures to one another.

Finally, we comment on the racial composition -- defined here as the percent of students who identify as Black, Latinx, or Native -- of private high schools visited by private colleges and universities in our sample. Figure \@ref(fig:ego-network-figure-privu) shows substantial variation across and within clusters in the geographic region, religious affiliation, and academic reputation of visited private schools. By contrast, the racial composition of visited private high schools was remarkably similar. On average, about one-third of visits were to schools where Black/Latinx/Native students comprised less than 10% of the student body, about 50% of visits were to schools where Black/Latinx/Native students comprised between 10% and 25% of the student body, and about 15% of visits were to schools where Black/Latinx/Native students comprsied more than 25% of the student body. Some colleges (e.g., Williams and Swarthmore) were better than others (e.g., Emory, Sewanee) at visiting schools where Black/Latinx/Native students comprised more than 50% of the population, but these visits to very diverse private schools represented a small share of all visits to private schools. The overall results suggest that when private colleges and universities visit private high schools, they are usually visiting schools that enroll few Black, Latinx, or Native students.

[HOW DOES RACIAL COMPOSITION OF VISITED PRIVATE HIGH SCHOOLS COMPARE TO RACIAL COMPOSITION OF ALL PRIVATE HIGH SCHOOLS? CUT/REVISE TEXT DEPENDING ON ANSWER TO THIS QUESTION] See Figure \@ref(fig:privhs-universe). Private high schools tend to have a higher share of White students than public high schools. Moreover, the private high schools that were visited by private colleges and universities in our sample had a lower share of Black/Latinx/Native students, on average, than the population of private high schools. Our findings suggest that the low number of Black/Latinx/Native students at private colleges and universities may be driven by thse institutions explicitly recruiting from private high schools that enroll few Black/Latinx/Native students.

[NEED TRANSITION TO FOCUS NOW ON HIGH SCHOOLS] Which private high schools received the most visits from private colleges and universities in our sample? Table \@ref(tab:hs-print-privu) shows characteristics of the 20 private high schools that were visited by the most private colleges and universities. The last two columns of Table \@ref(tab:hs-print-privu) are "degree" and "strength," which are related measures of vertex "centrality." Measures of centrality attempt to identify the most important or influential actors in a social network. "Degree" is defined as the number of edges direclty connected to a vertex. In Table \@ref(tab:hs-print-privu), the degree of a high school is the number of private colleges and universities in our sample that visited. The concept "strength" is generally defined as a weighted measure of degree. In our study, colleges often made more than one visit to the same high school in 2017. Thus, Table \@ref(tab:hs-print-privu) defines strength as the total number of visits from private colleges and universities in our sample.

The Kent Denver School, the top ranked school in \@ref(tab:hs-print-privu),was visited by 23 of the 26 private colleges and universities in our sample. Choate Rosemary Hall was visited by 21 colleges and universities, but these 21 institutions made 34 total visits. Most schools in Table \@ref(tab:hs-print-privu) are mostly nonsectarian. Although all received a Niche grade of "A+", they are not uniformly in the upper-echelon of schools ranked in the top 100 nationwide. With the exception of St. Ignatius Prep in Chicago and the Maret School in Washington DC, students who identify as Black, Latinx, or Native comprise less than 20% of all students.

Table \@ref(tab:hs-table-privu-agg) shows the characteristics of private high schools by degree band (e.g., visits from 15+, 10-14, etc). Schools that received visits from 4 or fewer private colleges and universities enrolled a greater share of Black/Latinx/Native students. These schools were also disproportionately Catholic schools and were more likely to recieve a grade of "A" or lower from Niche.

Whereas Tables \@ref(tab:hs-print-privu) and \@ref(tab:hs-table-privu-agg) examine centrality, they do not assess the extent to which high schools cluster into groups, with some high schools visited by nearly all colleges and universities and other groups of high schools only receiving visits from a particular subset of colleges and universities. [THIS IS WHERE THE DIRECT VS. INDIRECT CONNECTION ANALYSIS RECOMMENDED BY RUSS WOULD GO. WE MIGHT ADD THIS TO BURD CHAPTER OR DO AFTER BURD CHAPTER SUBMITTED]

## Visits by Public Research Universities

```{r, include = FALSE, eval = FALSE}
df <- readRDS('./assets/tables/table_ego_pubu.RDS')
df %>% glimpse()
df %>% select(ID,University,Cluster,Rank,total, in_state, out_state) %>% 
  arrange(Cluster,desc(total)) %>% print(n=30)
```

Figure \@ref(fig:plot-1mode-pubu) plots the one-mode network of visits by 17 public research universities, with vertex color determined by hierarchical cluster analysis specified to create $k = 4$ clusters. Accompanying Figure \@ref(fig:plot-1mode-pubu) and created from the respective ego networks of each university is Figure \@ref(fig:ego-network-figure-pubu), which shows the characteristics of private high schools visited by these public reearch universities.

We briefly sketch broad findings before discussing each cluster. First, for must public research universities in our sample, the vast majority of private school visits were to out-of-state schools [WHICH FIGURE/TABLE SHOWS THIS], which is consistent with the idea that public research universities pursue students from affluent households who can afford nonresident tuition price. Second, Figure \@ref(fig:ego-network-figure-pubu) indicates that public research universities tended to visit private high schools with weaker academic reputation than the set of schools visited by private colleges and universities [LATER, SAY GOING AFTER MONEY WITHOUT ACADEMIC REP]. Third, despite being nonsectarian, most public research universities visited a higher share of private high schools that were Catholic or Chrisitan than those visited by private colleges and universities. Fourth, racial composition of schools visited by public research universities was similar to the set of schools visited by private colleges and universities. A few public research universities appear (e.g., UC Irvine, UC Riverside) appear to visit a more racially diverse set of schools, but these institutions visited very few private schools.

<!--
Truly singular in the realm of recruiting, the University of Alabama is a cluster unto itself
-->

The University of Alabama is a cluster unto itself and is centrally located in the network graph in Figure \@ref(fig:plot-1mode-pubu). In 2017, the University of Alabama made 985 visits to U.S. private high schools. This number was more than double the 486 visits made by the University of South Carolina, which ranked second amongst public research universities in terms of number of private school visits, and substantially more than the 687 visits made by Notre Dame, which ranked first amongst private colleges and universities. Drawing from Figure \@ref(fig:ego-network-figure-pubu), more than 50% of visits by the University of Alabama were in the South. With respect to religious affiliation, nearly 50% of visits were to Catholic schools and about 25% of visits were to Christian schools. In contrast to the majority of private colleges and universities in our sample, the University of Alabama made nearly 50% of its visits at schools that received a grade of less than "A+" by Niche. At the same time 80% of visited schools enrolled fewer than 25% of students who identified as Black, Latinx, or Native.

A second cluster, located nearly as centrally in Figure \@ref(fig:plot-1mode-pubu) as the University of Alabama, consists of the University of South Carolina (486 visits) and the University of Colorado-Boulder (379 visits). These institutions ranked second and third amongst public universities in our sample in terms of number of visits.  Geographically, both universities recruiting nationally but concentrated on some regions more than others. For both universities, the Northeast and Midwest accounted for about 25% of visits.  The University of South Carolina concentrated visits in its home region, the South. CU-Boulder made X% of visits in its home region, the West, and made X% of visits in the South.  Both universities made more than 60% of visits at Catholic or Christian affiliated schools. In terms of academic reputation, CU-Boulder visited a slightly stronger set of schools than the University of South Carolina and both universities visited a stronger set of schools than the University of Alabama.

Third, the largest cluster -- at nine universities -- and the most diverse in terms of geographic location and academic reputation consists of three universities from the Northeast (Rutgers, UMass Amherst, University of Pittsburgh), two Midwest universities (University of Cincinnati, University of Kansas), two universities in the South (University of Arkansas, University of Georgia), and two universities from the West (UC Berkeley, UC San Diego).  Located between the center and periphery in Figure \@ref(fig:plot-1mode-pubu), these universities made a moderate number of private school visits, ranging from 197 by UC-San Diego to 302 and 303, respectively, by the University of Georgia and Rutgers.

These universities differ in the extent to which visits were concentrated in their home geographic region. At the high end, the University of Georgia -- 47th in YYYY US News rankings -- and the University of Arkansas -- 160th in US News rankings -- are located in the South and made at least 75% of visits in the South. The remaining seven universities generally made more visits in their home geographic than other regions but while also making substantial visits outside their home region, particularly in the South. For example, three highly ranked universities -- University of Pittsburgh (ranked 58), UC-Berkeley (ranked 22), and UC-San Diego  (ranked 35) each made a substantial share of visits in the South. In contrast to private colleges and universities, several public research universities in this cluster largely ignored particular geographic regions altogether. For example, the University of Kansas and the University of Arkansas did not visit the Northeast. The University of Georgia, UC-San Diego, and UMass Amherst largely ignored the MidWest, while the University of Pittsburgh did not visit the West.

>>>>

---
title: "Recruiting Chapter"
author: 
  - Ozan Jaquette
  - Crystal Han
  - Irma Castaneda
bibliography: ./../other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./../other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \floatsetup{capposition=top}
---

@@

Prior to the pandemic, admissions counselors from colleges and universities traveled to local high schools each year in order to meet prospective students and maintain relationships with guidance counselors at feeder schools [@RN3519]. @RN4721 analyzed off-campus recruitiving visits made in 2017 by 15 public research universities. Surprisingly, 12 of the 15 universities made more visits to *out-of-state* high schools than *in-state* high schools. These out-of-state visits focused on public schools in affluent, predominantly white communities. Additionally, most public research universities in our sample made a disproportionate number of visits to out-of-state private high schools. For the University of California-Berkeley, for example, 155 (43%) of the 360 out-of-state high school visits were to private high schools. For the University of Colorado-Boulder, 321 (36%) of the 901 out-of-state high school visits were to private high schools.

Whereas @RN4721 primarily analyzed visits to public high schools, this chapter focuses on visits to private high schools. More specifically, we investigate how visits to private high schools by public research universities compare to visits to private high schools by selective private universities.

## Social mobility, social reproduction, and privatization

The sociologist Max Weber is credited for the insight that education has a "dual character," on one hand, serving *social mobility* by redistributing access to privileged positions and, on the other, serving *social reproduction* by granting privileged families disproportionate access to the educational institutions that confer pathways to privileged positions [@RN2377].

The traditional, albeit mythologized, public good mission of public research universities is tied to social mobility and development of the state [@RN2269; @RN921]. These institutions are designated the unique responsibility of educating the future civic, professional, and business leaders of the state. Quoting 19th century University of Michigan President James Angell, these institutions contribute to social mobility by providing “an uncommon education for the common man” [as cited in @RN3608, p. 279] who could not afford tuition at elite private institutions.

By contrast, the traditional mission of selective private colleges and universities is more squarely associated with reproduction. Historically, these institutions depend on patronage from philanthropic interests -- in connection with business interests and regligious congregations -- and depend on patronage from wealthy households who provide tuition revenue and donations [@RN4517]. In turn, selective private institutions contribute to social reproduction by providing privileged access -- and the opportunities it confers -- to their wealthy patrons.

Recruiting visits to private high schools are entirely consistent with the traditional social reproduction mission of selective private universities, regardless of whether the university is targeting a particular high school because of its academic reputation, religious affiliation, or the affluence of its student body. By contrast, visits to private high schools by public research universities -- and particularly to out-of-state private high schools -- particularly out-of-state schools -- are antithetical are antithetical to the public good mission of prioritizing meritorious state residents, particularly from humble means.

We argue that an analysis of recruiting visits to private high schools can contribute novel insights about the "privatization" of public research universities to the interdisciplinary literature on privatization in the field of higher education research. One strand within this literature conceptualizes privatization as a process whereby declines in government funding cause public institutions to grow alternative revenue sources, particularly revenue from tuition and research commercialization [@RN4543@; @RN2101]. Of relevance to this chapter, public research universities responded to declines in state appropriations by growing enrollment from nonresident students [@RN3753], who typically pay two- to three-times more than resident students. Another strand within the privatization literature conceptualizes privatization as a process whereby public institutions become more similar -- in terms of characteristics or behavior -- to for-profit or private non-profit organizations [@RN4543]. A weakness of this literature is that few empirical studies analyze public institutions on a behavior associated with private institutions.

This chapter compares public research universities to selective private universities on a behavior -- recruiting visits to private high schools -- that is consistent with the traditional mission of selective private institutions and antithetical to the traditional mission of public research universities. Empirical analyses are organized around the following three research questions:

1. How does the scale of visits to private high schools by public research universities compare to the scale of visits to private high schools by private universities?
1. How do the characteristics of private high schools visited by public research universities compare to the characteristics of private high schools visited by private universities?
1. To what extent do public research universities and private universities visit overlapping sets of private high schools?

# What do we know about off-campus recruiting visits

@@

ADD TABLE OF UNIVERSITY CHARACTERISTICS HERE
